---
title: "Project3"
author: "Jiaming Qu"
date: "11/17/2019"
output: html_document
---

## Background Introduction
The data set is kindly provided by Urban Ministries of Druham, which records their assistance and help to people each year. Different from Project 1 which includes services in many aspects, this project focuses on the shelter service with detailed information of clients and service records. In this project, I am going to explore how the shelter services influence clients and find the hidden trends.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(ggplot2)
library(tidyverse)
df <- read.csv("/Users/jiamingqu/Desktop/BIOS 611/project/bios611-projects-fall-2019-arthur0804/project3/data/cleaned/client_data_merged.tsv", sep = "\t")
```

## Research Question
### Section 1. Deomographic data analysis of our clients.
#### 1.1 How old are our clients when they come to the shelter?
```{r, echo=FALSE, message=FALSE, warning=FALSE}
# plot
ggplot(df, aes(x=Client.Age.at.Entry)) +  
    geom_histogram(binwidth = 5, color = "red", fill = "green", alpha = 0.4) + 
    labs(title ="Distribution of age when coming the shelter", x = "Age", y = "Count") +
    theme_grey()
```

It shows that people's ages when coming to the shelter nearly follows a normal distribution, with centering between 40 to 60 years old.

#### 1.2 What about gender and race?
```{r, echo=FALSE, message=FALSE, warning=FALSE}
gender <- df %>% filter((Client.Gender == "Female") | (Client.Gender == "Male"))
ggplot(gender, aes(fill=Client.Primary.Race, x=Client.Gender)) + 
  geom_bar(position="stack") + 
  labs(title ="Distribution of gender and race", x = "Gender", y = "Count") +
  theme_grey()
```

It shows that there are more male clients than female clients (4000 vs near 1000). Also we could see the most clients are Black or African American, followed by White.


## Behavior Analysis
### 2.1 How long does our clients stay in the shelter everytime when they come?
```{r, echo=FALSE, message=FALSE, warning=FALSE}
# filter invalid length
df_length_filtered <- subset(df, length.of.stay >= 0)
# plot
ggplot(df_length_filtered, aes(x=length.of.stay)) +  
    geom_histogram(binwidth = 30, color = "red", fill = "green", alpha = 0.4) + 
    labs(title ="Distribution of stay in the shelter", x = "Days", y = "Count") + 
    scale_x_continuous(limits = c(0, 750)) + 
    scale_y_continuous(limits = c(0, 1500)) + 
    theme_grey()
```

It shows that most people stay under 100 days in the shelter with few up to 250 days everytime when they come to the shelter.

### 2.2 Are there any correlations between people's age when coming to the shelter and their length of stay?
```{r, echo=FALSE, message=FALSE, warning=FALSE}
# plot
ggplot(df, aes(x=Client.Age.at.Entry, y=length.of.stay)) +  
    geom_point() + 
    labs(title ="Correlation between age and length of stay", x = "Age", y = "Days") +
    theme_grey()
```

We do not see a positive correlation between age and length of stay, my hypothesis that the older people are when they come to the shelter, the longer they will stay is proved to be wrong.

### 2.3 Do our clients come back to our shelter after they leave?
```{r, echo=FALSE, message=FALSE, warning=FALSE}
# get frequency
df_frequency <- as.data.frame(table(df['Client.ID']))
# plot
ggplot(df_frequency, aes(x=Freq)) +  
    geom_histogram(binwidth = 1, color = "red", fill = "green", alpha = 0.4) + 
    labs(title ="Distribution of frequency coming to the shelter", x = "Frequency", y = "Count") + 
    scale_x_continuous(limits = c(0, 20)) + 
    theme_grey()
```

It shows that most people only come under 3 times to our shelter, with 1 time as the most. Therefore we might prepare for someone's coming back.

### 2.4 What month do clients like to come?
```{r, echo=FALSE, message=FALSE, warning=FALSE}
# conver date time
df$Entry.Date <- as.Date(df$Entry.Date, "%Y-%m-%d")
df$year <- as.numeric(format(df$Entry.Date,'%Y'))
df$month <- as.numeric(format(df$Entry.Date,'%m'))

# dataframe of new client counts by year
df_avg_month <- data.frame(matrix(ncol = 2, nrow = 0))
x <- c("Clients", "Month")
colnames(df_avg_month) <- x

month <- c(1:12)
for (m in month){
  # extract the record for this year
  df_sum_by_month <- filter(df, df$month == m)
  s  = nrow(df_sum_by_month)
  s = s / 7
  df_avg_month[nrow(df_avg_month) + 1,] = c(s, m)
}

ggplot(df_avg_month, aes(x=Month, y = Clients)) + 
  geom_histogram(stat="identity") +
  labs(title ="Amount of clients by month (average of years)", 
       x = "Month", y = "Count") +
  scale_x_continuous(breaks = c(1, 3, 5, 7, 9, 11)) + 
  theme_grey()
```


This indicates that on average of these years, there is no a particular month when our clients like to come. Three seasons have the most records, which are March, May and August. Thus, we need to parepare for these seasons and also all year round. 

## 3. Other Information
### 3.1 How does the trend of people's coming grow in general?
```{r, echo=FALSE, message=FALSE, warning=FALSE}
# dataframe of client counts by year
df_clients_year <- df %>% group_by(year) %>% summarise(Clients = n())

# dataframe of new client counts by year
df_new_clients_year <- data.frame(matrix(ncol = 2, nrow = 0))
x <- c("NewClients", "year")
colnames(df_new_clients_year) <- x
year <- c(2012, 2014, 2015, 2016, 2017, 2018, 2019)

# loop over each year
for (i in year){
    # extract the record for this year
    df_clients_this_year <- filter(df, df$year == i)
    # get the unique client numbers of this year
    count_this_year <- unique(df_clients_this_year$Client.ID)
    # extract the record for past years
    df_clients_past_year <- filter(df, df$year < i)
    # get the unique client numbers of this year
    count_past_year <- unique(df_clients_past_year$Client.ID)
    # append count of new clients to the dataframe
    new_clients_count = length(setdiff(count_this_year, count_past_year))
    df_new_clients_year[nrow(df_new_clients_year) + 1,] = c(new_clients_count, i)
}

# combine two dataframes
df_clinets <- inner_join(df_clients_year, df_new_clients_year, by = "year")

# reshape
df_clinets <- df_clinets %>% gather(type, count, c(Clients, NewClients)) 

# plot
ggplot(df_clinets, aes(x=year, y = count, color = type)) + 
    geom_line() +
    labs(title ="Amount of clients and new clients by year", 
         x = "Year", y = "Count", color = "Amount Of") +
    theme_grey()
```

This figure shows that apart from 2012-2014, the general trends of clients and new clients coming to the shelter are decreasing.

### 3.2 Where do our clients go after they leave our shelter?
```{r, echo=FALSE, message=FALSE, warning=FALSE}
leaving_df <- df %>% group_by(Destination)  %>% summarise(count = n()) %>% arrange(desc(count))
leaving_df <- leaving_df[1:5,]

ggplot(leaving_df, aes(x=Destination, y = count)) + 
  geom_histogram(stat="identity") +
  labs(title ="Destination of our clients after leaving the shelter", 
       x = "Destination", y = "Count") +
  theme_grey() + 
  theme(axis.text.x = element_text(angle = 25, hjust = 1))
```

From the figure we could see the most impossible destination for our clients are staying or living with friends, followd by with family with exception of "No Interview" which is the 2nd place.

### 3.3 What are the reasons that make our clients leave the shelter?
```{r, echo=FALSE, message=FALSE, warning=FALSE}
leaving_df2 <- df %>% group_by(Reason.for.Leaving)  %>% summarise(count = n()) %>% arrange(desc(count))
leaving_df2 <- leaving_df2[1:5,]

ggplot(leaving_df2, aes(x=Reason.for.Leaving, y = count)) + 
  geom_histogram(stat="identity") +
  labs(title ="Reasons of our clients' leaving the shelter", 
       x = "Reasons for leaving", y = "Count") +
  theme_grey() + 
  theme(axis.text.x = element_text(angle = 25, hjust = 1))
```

From the figure we could see the most impossible reasons for leaving are the completion of the program or left for housing, with exception of "Other" and "Unknown" which are the 1st and 2nd places.

### 3.4 Are there any correlations between clients' destination and the reasons?
```{r, echo=FALSE, message=FALSE, warning=FALSE}
# filter empty data
df <- df %>% filter(df$Destination != "")
df <- df %>% filter(df$Reason.for.Leaving != "")

# chi-squre test
t <- table(df$Destination, df$Reason.for.Leaving)
chisq <- chisq.test(t)
result <- round(chisq$residuals, 3)
result_df <- as.data.frame.matrix(result)

# store the result
df_cor <- data.frame(matrix(ncol = 3, nrow = 0))
x <- c("Destination", "Reason", "Value")
colnames(df_cor) <- x
row_names <- rownames(result_df)
col_names <- colnames(result_df)
for(i in 1:length(row_names)){
    for(j in 1:length(col_names)){
        value <- result_df[i,j]
        rowname <-row_names[i]
        colname <-col_names[j]
        df_cor[nrow(df_cor) + 1,] = c(rowname, colname, value)
    }
}

df_cor$Value <- as.numeric(df_cor$Value)
```

We use the Chi-squre test to test the correlation between two categorical data.

Top 10 positive correlation:
```{r, echo=FALSE, message=FALSE, warning=FALSE}
df_sorted_desc <- df_cor %>% arrange(desc(df_cor$Value))
for (i in c(1:10)){
  d = df_sorted_desc[i,1]
  r = df_sorted_desc[i,2]
  print(paste("Destination:",d, " + Reason: ", r))
}
```



